### 2018-08-25 introduction.tex
Conflicts resolved (hopefully)

\section{Introduction}
% State the objectives of the work and provide an adequate background, avoiding
% a detailed literature survey or a summary of the results.
The \gls{MSR} is an advanced nuclear reactor developed at \gls{ORNL} 
in the 1950s and operated in the 1960s. More recently, the \gls{GIF} included 
\glspl{MSR} among the six advanced reactor concepts promising for further 
research and development.
\glspl{MSR} offer significant improvements ``in the four broad areas of 
sustainability, economics, safety and reliability, and proliferation resistance 
and physical protection" \cite{doe_technology_2002}. To achieve the goals 
formulated by the GIF, \glspl{MSR} attempt to simplify the reactor core and 
improve inherent safety by using liquid fuel.

 In the thermal spectrum \gls{MSR}, fluorides of fissile and/or fertile 
 materials (i.e. UF$_4$, ThF$_4$,  PuF$_3$, TRU\footnote{ Transuranic 
 elements}F$_3$) combine with carrier salts to form a liquid fuel that 
 circulates in a loop-type primary circuit \cite{haubenreich_experience_1970}.  
 Immediate advantages over traditional, solid-fueled, reactors include 
 near-atmospheric pressure in the primary loop, 
 relatively high coolant temperature, outstanding neutron economy, improved 
 safety parameters, reduced fuel preprocessing, and the ability to 
 continuously remove fission products and add fissile and/or fertile elements 
 \cite{leblanc_molten_2010}. The thorium-fueled \gls{MSBR} was developed in the 
 early 1970s by \gls{ORNL} to explore the promise of the thorium 
 fuel cycle, which uses natural thorium instead of enriched uranium. With 
 continuous fuel reprocessing, the \gls{MSBR}
 realizes the advantages of the thorium fuel cycle because the $^{233}$U bred from 
 $^{232}$Th is almost instantly \footnote{\space $^{232}$Th transmutes into 
 $^{233}$Th after capturing a neutron. Next, this isotope decays to $^{233}$Pa 
 ($\tau_{1/2}$=21.83m), which finally decays to $^{233}$U 
 ($\tau_{1/2}$=26.967d).} recycled back to the core 
 \cite{betzler_modeling_2016}. The chosen fuel salt, 
 LiF-BeF$_2$-ThF$_4$-UF$_4$, has a $499^\circ$C melting point, a low vapor 
 pressure at operating temperatures, 
 and good flow and heat transfer properties \cite{robertson_conceptual_1971}. 
 In the matter of nuclear fuel cycle, the thorium cycle produces a reduced 
 quantity of plutonium and \glspl{MA} compared to the traditional uranium fuel 
 cycle. Finally, the \gls{MSR} also could be employed as a converter reactor 
 for transmutation of spent fuel from current \glspl{LWR}.

@@

References \cite{doligez_coupled_2014,heuer_towards_2014,sheu_depletion_2013,
aufiero_extended_2013} provide some form of reactivity control, and methods 
\cite{doligez_coupled_2014,heuer_towards_2014,aufiero_extended_2013,
ahmad_neutronics_2015, park_whole_2015, jeong_equilibrium_2016,rykhlevskii_online_2017,
nuttin_potential_2005} use a set of all nuclides in depletion calculations.

Liquid-fueled \gls{MSR} designs have online separations and feeds, where material 
is moved to or from the core at specific time steps (batch) or continuously. In 
batch-wise approach the burn-up simulation stops at a given time and restart with
a new liquid fuel composition (after removal of discarded materials and addition 
of fissile/fertile materials). \gls{ORNL} researchers have developed ChemTriton, 
a Python-based script for SCALE/TRITON which the batch-wise to simulate a continuous 
reprocessing. This tool models salt treatment, separations, discharge, and refill 
using a unit-cell \gls{MSR} SCALE/TRITON depletion simulation over small time steps 
to simulate continuous reprocessing and deplete the fuel salt \cite{powers_new_2013}. 
Methods listed in references \cite{fiorina_investigation_2013,sheu_depletion_2013,
park_whole_2015, jeong_equilibrium_2016,powers_inventory_2014,betzler_molten_2017,
rykhlevskii_online_2017} and current work employ a batch-wise approach.

Accounting for a continuous removal or addition is more difficult because it requires
adding a term to the Bateman equations. In SCALE \cite{bowman_scale_2011}, 
ORIGEN \cite{gauld_isotopic_2011} solves a set of Bateman equations using 
spectrum-averaged fluxes and cross sections generated from a deterministic transport 
calculation. Approaches listed in references \cite{doligez_coupled_2014,heuer_towards_2014,
aufiero_extended_2013,nuttin_potential_2005} model true continuous feeds and removals.

Thorium-fueled \gls{MSBR}-like reactors similar to the one in this work are described in 
\cite{park_whole_2015, jeong_equilibrium_2016,powers_new_2013,powers_inventory_2014, 
betzler_molten_2017,rykhlevskii_online_2017,nuttin_potential_2005}. Nevertheless, most of 
these efforts considered only simplified unit-cell geometry because depletion computations 
for a many-year fuel cycle are computationally expesive even for simple models.

Nuttin \emph{et al.} broke up reactor core geometry into three \gls{MCNP} cells: 
one for salt channels, one for the salt plena above and below the core and the 
last cell for the annulus, consequently, the two-region reactor core was 
approximated by one region with averaged fuel/moderator ratio 
\cite{nuttin_potential_2005}.  A similar approach was used by Powers \emph{et 
al.}, Betzler \emph{et al.}, and Jeong \emph{et al.} 
\cite{powers_new_2013,powers_inventory_2014,betzler_modeling_2016, 
betzler_molten_2017, jeong_development_2014, jeong_equilibrium_2016}. This 
approach misrepresents the two-region breeder reactor concept. The unit-cell or 
one-region models may produce reliable results for homogeneous reactor cores 
(i.e. \gls{MSFR}, \gls{MOSART}) or for one-region single-fluid reactor designs 
(i.e. \gls{MSRE}). However, a two-region \gls{MSBR} must be simulated using a whole-core 
model to capture different neutron transport characteristics in the inner and 
outer regions of the core. In particular, most fissions happens in the inner 
region while breeding occurs in the outer zone.

>>>>

\section{Introduction}
% State the objectives of the work and provide an adequate background, avoiding
% a detailed literature survey or a summary of the results.
The \gls{MSR} is an advanced nuclear reactor which was developed at \gls{ORNL} 
in the 1950s and was operated in the 1960s. More recently, the \gls{GIF}  
included \glspl{MSR} among the six advanced reactor concepts promising for 
further research and development. 
\glspl{MSR} offer significant improvements ``in the four broad areas of 
sustainability, economics, safety and reliability, and proliferation resistance 
and physical protection" \cite{doe_technology_2002}. To achieve the goals 
formulated by the GIF, \glspl{MSR} attempt to simplify the reactor core and 
improve inherent safety by using liquid fuel.

 In the thermal spectrum \gls{MSR}, fluorides of fissile and/or fertile 
materials (i.e. UF$_4$, ThF$_4$,  PuF$_3$, TRU\footnote{ Transuranic 
elements}F$_3$) combine with carrier salts to form a liquid fuel that 
circulates in a loop-type primary circuit \cite{haubenreich_experience_1970}. 
Immediate advantages over traditional, solid-fueled, reactors include 
near-atmospheric pressure in the primary loop, 
relatively high coolant temperature, outstanding neutron economy, improved 
safety parameters, reduced fuel preprocessing, and the ability to continuously 
remove fission products and add fissile and/or fertile elements 
\cite{leblanc_molten_2010}. The thorium-fueled \gls{MSBR} was developed in the 
early 1970s by \gls{ORNL} specifically to explore the promise of the thorium 
fuel cycle, which uses natural thorium instead of enriched uranium. With 
continuous fuel reprocessing, the \gls{MSBR} realizes the advantages of the 
thorium fuel cycle because the $^{233}$U bred from 
$^{232}$Th is almost instantly \footnote{\space $^{232}$Th transmutes into 
$^{233}$Th after capturing a neutron. Next, this isotope decays to $^{233}$Pa 
($\tau_{1/2}$=21.83m), which finally decays to $^{233}$U 
($\tau_{1/2}$=26.967d).} recycled back to the core 
\cite{betzler_modeling_2016}. The chosen fuel salt, LiF-BeF$_2$-ThF$_4$-UF$_4$, has a 
melting point of $499^\circ$C, a low vapor pressure at operating temperatures, 
and good flow and heat transfer properties \cite{robertson_conceptual_1971}. In 
the matter of nuclear fuel cycle, the thorium cycle produces a reduced quantity 
of plutonium and \glspl{MA} compared to the traditional uranium fuel cycle. 
Finally, the \gls{MSR} also could be employed as a converter reactor for 
transmutation of spent fuel from current \glspl{LWR}.

@@

References 
\cite{doligez_coupled_2014,heuer_towards_2014,sheu_depletion_2013,aufiero_extended_2013} 
simulate some form of reactivity control, and methods 
\cite{doligez_coupled_2014,heuer_towards_2014,aufiero_extended_2013,ahmad_neutronics_2015, 
park_whole_2015, 
jeong_equilibrium_2016,rykhlevskii_online_2017,nuttin_potential_2005} use a set 
of all nuclides in depletion calculations.

Many liquid-fueled \gls{MSR} designs rely on online fuel processing in which  
material moves to and from the core at specific time steps (batch) or 
continuously. In batch-wise approach the burn-up simulation stops at a given 
time and restart with a new liquid fuel composition (after removal of discarded 
materials and addition of fissile/fertile materials). \gls{ORNL} researchers 
have developed ChemTriton, a Python-based script for SCALE/TRITON which uses the 
batch-wise approach to simulate a continuous reprocessing. ChemTriton models salt 
treatment, separations, discharge, and refill using a unit-cell \gls{MSR} 
SCALE/TRITON depletion simulation over small time steps to simulate continuous 
reprocessing and deplete the fuel salt \cite{powers_new_2013}. Methods listed in 
references \cite{fiorina_investigation_2013,sheu_depletion_2013,park_whole_2015, 
jeong_equilibrium_2016,powers_inventory_2014,betzler_molten_2017,rykhlevskii_online_2017} 
and current work employ a batch-wise approach.

Accounting for continuous removal or addition presents a greater challenge since it 
requires adding a term to the Bateman equations. In SCALE 
\cite{bowman_scale_2011}, ORIGEN \cite{gauld_isotopic_2011} solves a set of 
Bateman equations using spectrum-averaged fluxes and cross sections generated 
from a deterministic transport calculation. Approaches listed in references 
\cite{doligez_coupled_2014,heuer_towards_2014,aufiero_extended_2013,nuttin_potential_2005} 
model true continuous feeds and removals.

Thorium-fueled \gls{MSBR}-like reactors similar to the one in this work are 
described in \cite{park_whole_2015, 
jeong_equilibrium_2016,powers_new_2013,powers_inventory_2014, 
betzler_molten_2017,rykhlevskii_online_2017,nuttin_potential_2005}. 
Nevertheless, most of these efforts considered only simplified unit-cell 
geometry because depletion computations for a many-year fuel cycle are 
computationally expensive even for simple models.

Nuttin \emph{et al.} broke up reactor core geometry into three \gls{MCNP} cells: 
one for salt channels, one for the salt plena above and below the core and the 
last cell for the annulus, consequently, the two-region reactor core was 
approximated by one region with averaged fuel/moderator ratio 
\cite{nuttin_potential_2005}.  A similar approach was used by Powers \emph{et 
al.}, Betzler \emph{et al.}, and Jeong \emph{et al.} 
\cite{powers_new_2013,powers_inventory_2014,betzler_modeling_2016, 
betzler_molten_2017, jeong_development_2014, jeong_equilibrium_2016}. This approach 
misrepresent the two-region breeder reactor concept. The unit-cell or one-region 
models may produce reliable results for homogeneous reactor cores (i.e. 
\gls{MSFR}, \gls{MOSART}) or for one-region single-fluid reactor designs (i.e. 
\gls{MSRE}). However, a two-region \gls{MSBR} must be simulated using a whole-core 
model to capture different neutron transport characteristics in the inner and 
outer regions of the core. In particular, most fissions happens in the inner 
region while breeding occurs in the outer zone.

### 2018-08-25 results.tex
Conflicts resolved (hopefully)

\begin{figure}[ht!] 
  \centering
  \includegraphics[width=\textwidth]{keff.png}
  \caption{Effective multiplication factor dynamics for full-core \gls{MSBR} 
  model for a 60-year reactor operation. The confidence interval 
  $\sigma=\pm15pcm$ is shaded.}
  \label{fig:keff}
\end{figure}

@@

\subsection{Fuel salt composition dynamics}

The analysis of the fuel salt composition evolution provides more comprehensive 
information about the equilibrium state. Figure~\ref{fig:adens_eq} shows number 
density of major nuclides which have a strong influence on the reactor core 
physics. Concentration of $^{233}$U, $^{232}$Th, $^{233}$Pa, and $^{232}$Pa in 
fuel salt change insignificantly after approximately 2500 days of operation. 
Particularly, $^{233}$U number density fluctuates less than 0.8\% in the time 
interval from 16 to 20 years of operation, hence,a quasi-equilibrium state was 
achieved after 16 years of reactor operation.
\begin{figure}[ht!] % replace 't' with 'b' to 
  \centering
  \includegraphics[width=\textwidth]{major_isotopes_adens.png}
  \caption{Number density of major nuclides during 60 years of reactor 
  operation.}
  \label{fig:adens_eq}
\end{figure}
In contrast, a wide variety of nuclides, including fissile isotopes (e.g. 
$^{235}$U) and non-fissile strong absorbers (e.g. $^{234}$U), keep accumulating 
in the core. Figure~\ref{fig:fissile_short} demonstrates production of fissile 
isotopes in the core. In the end of the considered operational time, the core 
contains significant $^{235}$U ($\approx10^{-5}$ atom/b-cm), $^{239}$Pu 
($\approx5\times10^{-7}$ atom/b-cm), and $^{241}$Pu ($\approx 5\times10^{-7}$ 
atom/b-cm). Meanwhile, the equilibrium number density of the target fissile 
isotope $^{233}$U was approximately 7.97$\times10^{-5}$ atom/b-cm. Thus, 
production of new fissile materials in the core as well as $^{233}$U breeding 
make it possible to compensate for negative effects of strong absorber 
accumulation and keep the reactor critical.
\begin{figure}[htp!] % replace 't' with 'b' to \centering
  \includegraphics[width=\textwidth]{fissile_short.png}
  \caption{Number density of fissile in epithermal spectrum nuclides 
  accumulation during the reactor operation.}
  \label{fig:fissile_short}
\end{figure}

@@

Figure~\ref{fig:spectrum_zones} shows that zone I produced more thermal neutrons 
than zone II, corresponding to a majority of fissions occuring in the central part 
of the core. In the undermoderated zone II, the neutron energy spectrum is harder 
which leads to more capture of neutrons by $^{232}$Th and helps a achieve relatively 
high breeding ratio. Moreover, the (n,$\gamma$) resonance energy range in $^{232}$Th 
is from 10$^{-4}$ to 10$^{-2}$ MeV. Therefore, the moderator-to-fuel ratio for zone 
II was chosen to shift the neutron energy spectrum in this range. Furthermore, in the 
central core region (zone I), the neutron energy spectrum shifts to a harder spectrum 
over 20 years of reactor operation. In contrast, in the outer core region (zone II) a 
similar spectral shift takes place at a reduced scale. This resuls is in a good 
agreement with original ORNL report \cite{robertson_conceptual_1971} and most recent 
whole-core steady-state study \cite{park_whole_2015}.

@@

The effective multiplication factor of the core reduces gradually over 
operation time because the fissile material ($^{233}$U) continuously depletes 
from the fuel salt due to fission and, at the same time, fission products 
accumulates in the fuel salt.	Eventually, without fission products removal, 
the reactivity decreases to the subcritical state after approximately 500 and 
1300 days of operation for cases with no removal and volatile gases \& noble 
metals removal, respectively. The time when the simulated core reaches 
subcriticality ($k_{eff}<$1.0) for full-core model) called the core lifetime. 
Therefore, removing fission products provides with significant neutronic benefit 
because allows achieve longer core lifetime.
\begin{figure}[ht!] % replace 't' with 'b' to force it to 
  \centering
  \includegraphics[width=\textwidth]{keff_rem_cases.png} 
  \caption{Calculated effective multiplication factor for full-core \gls{MSBR} model with removal of different fission product groups over 10 years of operation.}
  \label{fig:fp_removal}
\end{figure}

>>>>

\begin{figure}[ht!] 
  \centering
  \includegraphics[width=\textwidth]{keff.png}
  \caption{Effective multiplication factor dynamics for full-core \gls{MSBR} 
  model for a 60-year reactor operation.}
  \label{fig:keff}
\end{figure}

@@

\subsection{Fuel salt composition dynamics}
The analysis of the fuel salt composition evolution provides more comprehensive 
information about the equilibrium state. Figure~\ref{fig:adens_eq} shows number 
density of major nuclides which have a strong influence on the reactor core 
physics. The concentration of $^{233}$U, $^{232}$Th, $^{233}$Pa, and $^{232}$Pa in 
fuel salt change insignificantly after approximately 2500 days of operation. 
Particularly, $^{233}$U number density fluctuates less than 0.8\% in the time 
interval from 16 to 20 years of operation. Hence, a quasi-equilibrium state was 
achieved after 16 years of reactor operation.
\begin{figure}[ht!] % replace 't' with 'b' to 
  \centering
  \includegraphics[width=\textwidth]{major_isotopes_adens.png}
  \caption{Number density of major nuclides during 60 years of reactor 
  operation.}
  \label{fig:adens_eq}
\end{figure}
In contrast, a wide variety of nuclides, including fissile isotopes (e.g. 
$^{235}$U) and non-fissile strong absorbers (e.g. $^{234}$U), keep accumulating 
in the core. Figure~\ref{fig:fissile_short} demonstrates production of fissile 
isotopes in the core. In the end of the considered operational time, the core 
contains significant $^{235}$U ($\approx10^{-5}$ atom/b-cm), $^{239}$Pu 
($\approx5\times10^{-7}$ atom/b-cm), and $^{241}$Pu ($\approx 5\times10^{-7}$ 
atom/b-cm). Meanwhile, the equilibrium number density of the target fissile 
isotope $^{233}$U was approximately 7.97$\times10^{-5}$ atom/b-cm. Thus, 
production of new fissile materials in the core as well as $^{233}$U breeding 
make it possible to compensate for negative effects of strong absorber 
accumulation and keep the reactor critical.
\begin{figure}[htp!] % replace 't' with 'b' to 
  \centering
  \includegraphics[width=\textwidth]{fissile_short.png}
  \caption{Number density of fissile in epithermal spectrum nuclides 
  accumulation during the reactor operation.}
  \label{fig:fissile_short}
\end{figure}

@@

Figure~\ref{fig:spectrum_zones} shows that zone I produced more thermal neutrons 
than zone II, corresponding to a majority of fissions occuring in the central part 
of the core. In the undermoderated zone II, the neutron energy spectrum is harder 
which leads to more capture of neutrons by $^{232}$Th and helps a achieve relatively 
high breeding ratio. Moreover, the (n,$\gamma$) resonance energy range in $^{232}$Th 
is from 10$^{-4}$ to 10$^{-2}$ MeV. Therefore, the moderator-to-fuel ratio for zone 
II was chosen to shift the neutron energy spectrum in this range. Furthermore, in the 
central core region (zone I), the neutron energy spectrum shifts to a harder spectrum 
over 20 years of reactor operation. Meanwhile, in the outer core region (zone II) a 
similar spectral shift takes place at a reduced scale. This results is in a good 
agreement with original ORNL report \cite{robertson_conceptual_1971} and most recent 
whole-core steady-state study \cite{park_whole_2015}.

@@

The effective multiplication factor of the core reduces gradually over 
operation time because the fissile material ($^{233}$U) continuously depletes 
from the fuel salt due to fission and, at the same time, fission products 
accumulates in the fuel salt.	Eventually, without fission products removal, 
the reactivity decreases to the subcritical state after approximately 500 and 
1300 days of operation for cases with no removal and volatile gases \& noble 
metals removal, respectively. The time when the simulated core reaches 
subcriticality ($k_{eff}<$1.0) for full-core model) called the core lifetime. 
Therefore, removing fission products provides with significant neutronic benefit 
because allows achieve longer core lifetime.
\begin{figure}[ht!] % replace 't' with 'b' to force it to 
  \centering
  \includegraphics[width=\textwidth]{keff_rem_cases.png} 
  \caption{Calculated effective multiplication factor for full-core \gls{MSBR} 
model with removal of various fission product groups over 10 years of 
operation.}
  \label{fig:fp_removal}
\end{figure}
